Obtaining a PhD is a long and challenging process, with many ups, downs, and intermediate results on the way there.
Nevertheless, all that formally matters are the dissertation and defense at the very end.
In this sense, I want to thank Prof. Peer Nowack and Prof. Ira Assent for serving as reviewers, Prof. Benjamin Schäfer for serving as the third examiner, and Prof. Klemens Böhm for the freedom and funding in the many years before.

To all PhD students reading this text:
Do not be fooled by the large number of people who are often mentioned in a dissertation's acknowledgments.
The PhD is, first and foremost, a personal project and essentially depends on your own motivation and work.
It can be a very lonely struggle from time to time.
If you do not advance your PhD project, nobody else will do.
However, other researchers and non-researchers can (and should) have a positive impact on your PhD's success, generally and in my personal case.

In the context of this dissertation, I want to thank Holger Trittenbach and Vadim Arzamasov for inspiration and advice on the topics that constitute my PhD research.
Further, my gratitude goes to Florian, Miki, Stas, and Tobias for proofreading.
Additional thanks go to all attendees of my defense and to those who made me a nice hat.

In the context of my work at KIT, several other persons made my PhD years more pleasant.
Elaheh, Markus, Pawel, and Tobias: Thanks for the successful collaborations and for showing me that collaborative research is much more fun than working alone, subject to having the right collaborators.
Edouard: Thanks for our joint undertaking to enhance the data science course.
Federico: Thanks for the sound teamwork regarding, but not limited to, the data science lab course.
Miki: Thanks for bringing cakes, games, and humor to the chair.
Pawel: Thanks for deep conversations on the PhD, research in general, and various other topics.
The rest of the chair: Thanks for also being there (participation trophy).
I particularly want to thank my fellow board-game and table-soccer players.
For legal reasons, I hereby declare that all playing took place during breaks and after work.

Finally, I want to thank my family and friends for their moral support, though they will not read this dissertation anyway.
(Honestly -- why should they?)
Special thanks go to my mum, whose systematic bread-slicing activities inspired me to investigate the problem of \textsc{Number Partitioning}, which kicked off my journey into complexity theory.
